Episodic memory: mental time travel or a quantum `memory wave' function?

### 2019-01-15 main.tex
added DuBrDava16 and MontEtal19 references

One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.  An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.  Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.  This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12, AlyEtal18}.  Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}. This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}. Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}.  These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.  In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.  Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.  First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).  Second, the studies suggest that members of this family of representations drifting at different rates become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.  These reactivations explain how our brains reactivate prior contexts when we remember the past.

@@

Unlike random word lists, naturalistic experiences contain \textit{event boundaries} whereby contextual cues change sharply from one moment to the next, implying a change in the current situation (e.g., shifting from the quiet peace of working on a manuscript in the early morning to the frenzied chaos of helping a newly awoken child get ready for school).  A number of studies (using non-random lists and more naturalistic stimuli) have found that the way we remember past experiences can be shaped by these event boundaries.  For example, controlling for elapsed time, memory is impaired for information that occurred prior to a change in the current event or situation~\citep[e.g., ][]{RadvCope06, SwalEtal09, SwalEtal11, EzzyDava11, MannEtal16}.  The temporal dynamics of contextual change also underlie how we judge the amount of time elapsed since a prior reference point~\citep{BlocReed78, SahaSmit14}.  These findings suggest that the rapid contextual or situational changes that define these event boundaries shape how we experience and remember, similarly to how spatial boundaries~\citep[e.g., environmental barriers;][]{McKeBuzs16, BrunEtal18} shape how we experience and remember spatial environments and layouts, or how conceptual boundaries~\citep[e.g., distinctions between semantic categories;][]{BrunEtal18} shape how we experience and remember conceptual information.

>>>>

One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.  An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.  Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.  This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12, AlyEtal18}.  Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}. This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}. Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}, supporting precise temporal recall~\citep{MontEtal19}.  These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.  In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.  Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.  First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).  Second, the studies suggest that members of this family of representations drifting at different rates become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.  These reactivations explain how our brains reactivate prior contexts when we remember the past.

@@

Unlike random word lists, naturalistic experiences contain \textit{event boundaries} whereby contextual cues change sharply from one moment to the next, implying a change in the current situation (e.g., shifting from the quiet peace of working on a manuscript in the early morning to the frenzied chaos of helping a newly awoken child get ready for school).  A number of studies (using non-random lists and more naturalistic stimuli) have found that the way we remember past experiences can be shaped by these event boundaries.  For example, controlling for elapsed time, memory is impaired for information that occurred prior to a change in the current event or situation~\citep[e.g., ][]{RadvCope06, SwalEtal09, SwalEtal11, EzzyDava11, MannEtal16}.  The temporal dynamics of contextual change also underlie how we judge the amount of time elapsed since a prior reference point~\citep{BlocReed78, SahaSmit14}.  These findings suggest that the rapid contextual or situational changes that define these event boundaries shape how we experience and remember~\citep{DuBrDava16}, similarly to how spatial boundaries~\citep[e.g., environmental barriers;][]{McKeBuzs16, BrunEtal18} shape how we experience and remember spatial environments and layouts, or how conceptual boundaries~\citep[e.g., distinctions between semantic categories;][]{BrunEtal18} shape how we experience and remember conceptual information.
